---
title: "DS152 Introduction to Data Science (2)"
subtitle: "Assignment 6: Unsupervised Learning"
author: "Prof. Niamh Cahill"
format: 
  html:
    embed-resources: true
    self-contained-math: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rpart)
library(rpart.plot)
library(class)
library(AppliedPredictiveModeling)
library(mclust)
```

## **Solutions** **\[Total = 20 Marks\]**


## Exercise

1.  Which of the following is not required for k-means clustering? 

```{=html}
<!-- -->
```


D)  All of the above are required. **\[2 Marks\]**

```{=html}
<!-- -->
```

2.  In k-means, k stands for 

```{=html}
<!-- -->
```


B)  Number of center points **\[2 Marks\]**


```{=html}
<!-- -->
```

3.  Which of the following can act as a possible termination condition for the k-means algorithm? 

```{=html}
<!-- -->
```

B)  When the assignment of observations to clusters does not change between iterations. **\[2 Marks\]**

```{=html}
<!-- -->
```

4.  For this analysis, we will work with the banknotes dataset to perform k-means clustering. First, we visualise the data by plotting Top vs Diagonal and colouring by Status (counterfeit, genuine) as shown below. Replace \[A\], \[B\], \[C\], \[D\] and \[E\] in the code below to create this plot. **\[5 Marks\]**


```{r, echo = TRUE}
  ggplot(banknote, aes(x = Top, y = Diagonal, colour = Status)) +
  geom_point()
```

\[A\] = aes
\[B\] = Top
\[C\] = Diagonal
\[D\] = Status
\[E\] = geom_point


5.  Now, let's perform k-means clustering. We will ignore the fact that the banknotes are classified as genuine or counterfeit, and work only with the variables `Top` and `Diagonal` to see if the k-means algorithm with k = 2 will cluster the genuine notes into a group and the counterfeit notes into a group. Firstly, we will create a new dataset that contains only the `Top` and `Diagonal` from the banknote dataset. Then we will run the kmeans algorithm with 2 clusters on this data. Replace \[A\],\[B\] and \[C\] in the code such that this objective can be achieved. **\[3 Marks\]**

```{r, echo = TRUE, include = FALSE}
banknote2 <- banknote %>% 
              select(Top, Diagonal)

kmean_res <- kmeans(banknote2, centers = 2)
```

\[A\] = select
\[B\] = kmeans
\[C\] = 2


6.  Next we want to visualise the results in the following way:

```{r, echo = TRUE,message=FALSE}
library(factoextra)
fviz_cluster(kmean_res , data = banknote2, geom = "point")
```

Replace \[A\], \[B\], \[C\] in the code below to create this plot. **\[3 Marks\]**

\[A\] = fviz_cluster
\[B\] = kmean_res
\[C\] = geom


7.  Based on comparing the plot in 3 with the plot in 1, did the k-means algorithm cluster all the genuine notes into the same cluster? Give a reason for your answer. **\[3 Marks\]**

  - No. It appears to have clustered one of the genuine notes incorrectly. However, this data point is in the middle of all the counterfeit datapoints so that is not surprising. 

