---
title: "DS152 Introduction to Data Science (2)"
subtitle: "Assignment 1: Tidyverse Recap"
author: "Prof. Niamh Cahill"
format: 
  html:
    embed-resources: true
    self-contained-math: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)


#https://rpubs.com/aruntrivedi/final_rr_pa2


nat_disaster_dat <- read_csv("https://www.dropbox.com/scl/fi/kepof8k55mzlefy7nzfrj/nat_disaster_dat.csv?rlkey=7nx78bcpnj6ufm0hdq91ke8np&raw=1")


```

## Instructions 
Answer all questions for continuous assessment. Due: **4pm on Friday, 28th February 2025**. Please write your answers in a Word file, save it as a **pdf document** and upload it to Moodle before the deadline. Alternatively, you may write down your answers using pen and paper, take photographs and upload a **pdf document** of the photos you've taken of your assignment. 



## Question

Synopsis: US Department of Commerce, National Oceanic & Atmospheric Administration provides information on fatalities and injuries associated with natural calamities/disasters across different US states.  

If you want to view the data in R you can run the following:

\scriptsize

```{r, eval = FALSE}
nat_disaster_dat <- read_csv("https://www.dropbox.com/scl/fi/kepof8k55mzlefy7nzfrj/nat_disaster_dat.csv?rlkey=7nx78bcpnj6ufm0hdq91ke8np&raw=1")
```


A glimpse if the data is shown below


```{r}
glimpse(nat_disaster_dat)
```

Note: 

  - `STATE` = US State
  - `EVTYPE` = event type
  - `FATALITIES` = \# of fatalities 
  - `INJURIES` = \# if injuries. 

(i) How many observations are in the dataset? 


(ii) Assume you want to create the summary below, which sums up the number of fatalities within each event type and then arranges by the total fatalities, in descending order. 

```{r, eval = TRUE, echo=FALSE}
nat_disaster_summary <- nat_disaster_dat %>% 
                          group_by(EVTYPE) %>% 
                          summarise(total_fatalities = sum(FATALITIES)) %>% 
                          arrange(desc(total_fatalities))
nat_disaster_summary
```

In the code below, what should [A], [B], [C] and [D] be replaced with to create this summary? 

```{r, eval = FALSE}
nat_disaster__summary <- nat_disaster_dat %>% 
                          group_by([A]) %>% 
                          [B](total_fatalities = [C])) %>% 
                          arrange([D](total_fatalities))
```

[A] = 

[B] = 

[C] = 

[D] = 

(iii) Now assume you want to filter the summarised data from (ii) so that you have a dataset with only events where the total fatalities are greater than 200, as shown.

```{r, eval = TRUE, echo=FALSE}
nat_disaster_filter <- nat_disaster_summary %>% 
                        filter(total_fatalities > 200)
nat_disaster_filter
```


In the code below, what should [A] and [B] be replaced with to create the filtered dataet? 


```{r, eval = FALSE}
nat_disaster_filter <- nat_disaster_summary %>% 
                        [A]([B])
```



[A] = 

[B] = 


(iv) Now assume you want to create the following barplot which shows the total fatalities per event type, using the filtered data from part (iii). 

```{r, eval = TRUE, echo=FALSE}
ggplot(nat_disaster_filter, aes(x = EVTYPE, y = total_fatalities)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("") +
  ylab("# fatalities")
```


In the code below, what should [A], [B] and [C] be replaced with to create the plot? 


```{r, eval = FALSE}
ggplot(nat_disaster_filter, aes(x = [A], y = [B])) +
  [C](stat="identity") +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("") +
  ylab("# fatalities")
```

[A] = 

[B] = 

[C] = 



