---
title: "R Notebook"
output: html_notebook
---

# Introduction to K-Means Clustering in R: Customer Segmentation

## 1. Overview of K-Means Clustering

### What is K-Means Clustering?

K-Means is an unsupervised machine learning algorithm used for partitioning a dataset into **K clusters** based on similarity. It minimizes the within-cluster sum of squares (WCSS) by iteratively refining cluster centroids.

### Why Use K-Means for Customer Segmentation?

Customer segmentation helps businesses identify distinct customer groups based on behavior, spending patterns, or demographics. This allows for targeted marketing and personalized services.

## 2. Dataset Description

We will use a **synthetic dataset** resembling customer spending behavior. The dataset contains: - `CustomerID`: Unique customer identifier - `Annual_Income`: Yearly income (in \$1000s) - `Spending_Score`: A score from 0 to 100 representing spending habits (higher means more spending)

## 3. Implementing K-Means in R

### Step 1: Load Libraries

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(cluster)
library(factoextra)
```

### Step 2: Load & Explore Data

```{r}
# Sample data: Simulated Customer Segmentation dataset
data <- data.frame(
  CustomerID = 1:200,
  Annual_Income = round(runif(200, 15, 120), 1),  # Income in $1000s
  Spending_Score = round(runif(200, 1, 100), 1)  # Spending Score (1-100)
)

# View first few rows
head(data)
```

### Step 3: Data Preprocessing

```{r}
# Select relevant features (dropping CustomerID)
data_scaled <- scale(data[, c("Annual_Income", "Spending_Score")])
```

### Step 4: Determine Optimal K (Elbow Method)

```{r}
set.seed(123) # For reproducibility
wcss <- vector()
for (k in 1:10) {
  kmeans_result <- kmeans(data_scaled, centers = k, nstart = 25)
  wcss[k] <- kmeans_result$tot.withinss
}

# Plot Elbow Method
plot(1:10, wcss, type = "b", pch = 19, frame = FALSE,
     xlab = "Number of Clusters (K)", ylab = "WCSS",
     main = "Elbow Method for Optimal K")
```

### Step 5: Apply K-Means Clustering

```{r}
# Apply K-Means with optimal K (let's assume K=5 from the Elbow Method)
kmeans_model <- kmeans(data_scaled, centers = 5, nstart = 25)
data$Cluster <- as.factor(kmeans_model$cluster)
```

### Step 6: Visualize the Clusters

```{r}
# Scatter plot of clusters
fviz_cluster(kmeans_model, data = data_scaled, geom = "point",
             ellipse.type = "norm", palette = "jco", ggtheme = theme_minimal())
```

## 4. Interpretation of Results

1.  **Identified Segments**: The customers are grouped into five distinct segments based on spending habits and income levels.
2.  **Cluster Characteristics**:
    -   High-income, high-spending customers (e.g., luxury spenders)
    -   High-income, low-spending customers (e.g., frugal earners)
    -   Low-income, high-spending customers (e.g., budget-conscious spenders)
    -   Low-income, low-spending customers (e.g., cost-sensitive buyers)
    -   Moderate-income, moderate-spending customers
3.  **Business Applications**: The segmentation results can inform targeted marketing strategies, loyalty programs, and pricing models.

## 5. Conclusion

-   **K-Means clustering** is a powerful tool for customer segmentation.
-   The **Elbow Method** helps select an appropriate number of clusters.
-   Visualization provides valuable insights into customer behavior.

------------------------------------------------------------------------

This example demonstrates how to implement K-Means clustering in R, interpret the results, and apply them to business scenarios.
