---
title: "DS152 Tutorial Sheet 8"
format: 
  html:
    embed-resources: true
    self-contained-math: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magick)
library(class)
library(rpart)
library(rpart.plot)

flowers_train <- readRDS("flowers_train.rds")
flowers_test  <- readRDS("flowers_test.rds")
```

## Exercise: Image Classification

We will classify images into three categories:

* Rose
* Sunflower
* Tulip

The raw images are in folders on the Rstudio server. The images are already preprocessed into datasets (also available on the Rstudio server) where:

* Each row represents one image
* Each column represents a pixel intensity
* The `label` column contains the correct class
* Training set: 24 images (`flowers_train`)
* Test set: 6 images (`flowers_test`)

We will compare two machine learning methods:

1. K-Nearest Neighbours (KNN)
2. Classification Tree

Both models will be trained on the training set and evaluated on the test set.



**(a)** Replace `[A]` and `[B]` in the code below to load and display the image `rose_1.png`.

```{r, eval = FALSE}
library(magick)

img <- image_[A]("flower_image_train/rose_1.png")
plot([B])
```

```{r, echo = FALSE}
library(magick)

img <- image_read("flower_image_train/rose_1.png")
plot(img)
```


**(b)** We now resize the image to 25 × 25 pixels. Replace `[A]` and `[B]` in the code below.

```{r, eval = FALSE}
img_scale <- image_[A]([B], "25x25!")
plot(img_scale)
```

```{r, echo = FALSE}
img_scale <- image_scale(img, "25x25!")
plot(img_scale)
```


**(c)** We now fit a KNN model using the training data and predict the test set labels.
Replace `[A]`, `[B]`, and `[C]` below.

```{r, eval = FALSE}
pred_knn <- [A](
  train = flowers_train %>% select(-[B]),
  test  = flowers_test %>% select(-[B]),
  cl    = flowers_train %>% pull([B]),
  k     = [C]
)
```

```{r, echo = FALSE}
pred_knn <- knn(
  train = flowers_train %>% select(-label),
  test  = flowers_test %>% select(-label),
  cl    = flowers_train %>% pull(label),
  k     = 5
)
```

Then the following code will produce the results data. 

```{r}
knn_res <- flowers_test %>%
  select(label) %>%
  mutate(pred_label = pred_knn)

knn_res
```



**(d)** We now fit a classification tree and visualise it. Replace `[A]` and `[B]` below.

```{r, eval = FALSE}
tree_model <- [A](label ~ ., 
                  data = flowers_train,
                  control = rpart.control(minsplit = 2))

rpart.plot([B])
```

```{r, echo = FALSE}
tree_model <- rpart(label ~ ., 
                    data = flowers_train,
                    control = rpart.control(minsplit = 2))

rpart.plot(tree_model)
```

Then the following code will produce the results data. 

```{r}
pred_tree <- predict(tree_model, newdata = flowers_test)

tree_res <- flowers_test %>%
  select(label) %>%
  mutate(pred_label = colnames(pred_tree)[max.col(pred_tree)])
```

**Notes**

  - The tree splits on individual pixel values.
  - This differs from KNN:
    - KNN uses similarity across all pixels
    - Trees use threshold rules on individual pixels




**(e)** We now compute confusion tables for both models. Calculate the accuracy in each case. 

```{r}
conf_knn <- table(knn_res)
conf_knn
```

```{r}
conf_tree <- table(tree_res)
conf_tree
```


```{r, include = FALSE, echo=FALSE}
N <- nrow(flowers_test)

acc_knn  <- sum(diag(conf_knn)) / N
acc_tree <- sum(diag(conf_tree)) / N

acc_knn
acc_tree
```


## Summary

**KNN**

  - Uses all pixel values
  - Measures overall similarity between images
  - Typically performs well on image data

**Classification Tree**

  - Uses only a small number of pixel-based rules
  - Easier to interpret
  - Often performs worse on raw image data because:
    - It does not capture relationships between pixels
    - Image data is high-dimensional and structured

