---
title: "DS152 Tutorial Sheet 1 with Solutions"
format: 
  html:
    embed-resources: true
    self-contained-math: true
editor: visual
---

## Exercise 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

untidy_tab1 <- table2
untidy_tab2a <- table4a
untidy_tab2b <- table4b
tidy_tab2b <- read_csv("https://www.dropbox.com/scl/fi/s0swvn1bnsyu1m04ewza5/tidy_tab2b.csv?rlkey=lwc63zi4o4g9st4787y1zo6jc&raw=1")

```

The table below displays the number of tuberculosis cases documented by the World Health Organization in Afghanistan, Brazil, and China between 1999 and 2000. The data contains values associated with four variables (country, year, cases, and population). The tidy format for the data is:

```{r, echo = FALSE}
table1
```

(a) The data table below organizes the same data in an untidy format.

```{r, echo = TRUE}
untidy_tab1
```

(i) What are the characteristics of this dataset that make it not tidy?

[The dataset is untidy because the observations are spread over multiple rows.]{style="background-color: lightblue;"}

(ii) In the code below, what should \[A\], \[B\], \[C\], \[D\] and \[E\] be replaced with to make this dataset tidy?

```{r}
tidy_tab1 <- untidy_tab1 %>% 
                pivot_wider(names_from = type ,
                            values_from = count)
tidy_tab1
```

[\[A\] = wider, \[B\] = names_from, \[C\] = type, \[D\] = values_from, \[E\] = count,]{style="background-color: lightblue;"}

(b) The data table below, which includes the cases data, is organized in an untidy format.

```{r, echo = TRUE}
untidy_tab2a
```

(i) What are the characteristics of this dataset that make it not tidy?

[Column names are values of the `year` variable.]{style="background-color: lightblue;"}

[The cases variable is spread over two columns.]{style="background-color: lightblue;"}

(ii) In the code below, what should \[A\], \[B\], \[C\], \[D\], \[E\] and \[F\] be replaced with to make this dataset tidy?

```{r}
tidy_tab2a <- untidy_tab2a %>%  
                pivot_longer(-country,
                              names_to = "year",
                              values_to = "cases")
tidy_tab2a
```

[\[A\] = longer, \[B\] = country, \[C\] = names_to, \[D\] = "year" (note "" around year is important here because you are creating a new variable), \[E\] = values_to, \[F\] = "cases" (note as with year, "" around cases is important here)]{style="background-color: lightblue;"}

(c) Once `tidy_tab2a` has been created you should make the `year` variable numeric. Replace \[G\] in the code below to achieve this.

```{r}
tidy_tab2a <- tidy_tab2a %>% mutate(year = as.numeric(year))
```

[\[G\] = as.numeric]{style="background-color: lightblue;"}

(d) The data table below includes the population data organized in a tidy format (note: China has an additional year of data):

```{r, echo = TRUE}
tidy_tab2b
```

Assume you want to join `tidy_tab2a` and `tidy_tab2b` in order to produce `table1`. Would you use `left_join` or `right_join`? Give a reason for your answer.

```{r, echo = TRUE}
left_join(tidy_tab2a,tidy_tab2b)
```

[`right_join` would include the additional year for China with an NA for the cases data and so would not produce `table1`.]{style="background-color: lightblue;"}

## Exercise 2

It is thought that the growth of a particular type of insect can be predicted by the temperature at which it is living in. In an experiment, eight insects of the same weight were given living temperatures ranging between 2\textsuperscript{o}C and 16\textsuperscript{o}C and their weight gain was recorded after three days. The data have been put into the following table.

```{r}
insect_id <- c(1, 2,3,4,5,6,7,8)
temperature <- c(2 ,4 ,6 ,8 ,10,12,14,16)
weight_gain <- c(1.21,0.96,1.31,1.52,1.41,1.43,1.87,1.67)

insect_dat <- tibble(insect_id, temperature, weight_gain)
insect_dat
```

(a) The correlation between temperature and weight gain is 0.84. What code would you use to do this calculation in R?

```{r}
cor(insect_dat$temperature, insect_dat$weight_gain)
```

(b) Interpret the correlation result.

[There is a strong positive correlation between temperature and weight gain.]{style="background-color: lightblue;"}

(c) What does it mean to say that two variables are negatively correlated?

[As one variable increases in value the other decreases.]{style="background-color: lightblue;"}

(d) Why should you inspect a scatter plot of the data even though you have information on the correlation between two variables?

```{r}
ggplot(insect_dat, aes(x = temperature, y = weight_gain)) +
  geom_point()
```

[Inspecting a scatter plot of the data is valuable for several reasons, even if you already know the correlation between two variables:]{style="background-color: lightblue;"}

[1. **Visual Confirmation**: While correlation coefficients provide a numerical measure of the strength and direction of the relationship between two variables, a scatter plot allows you to visually confirm this relationship. Sometimes, the correlation coefficient might indicate a weak relationship, but upon visual inspection, you might notice a clear pattern or non-linear trend in the data that was not apparent from the correlation alone.]{style="background-color: lightblue;"}

[2. **Outlier Detection**: Scatter plots help in identifying outliers, which are data points that deviate significantly from the overall pattern of the data. Outliers can have a disproportionate influence on the correlation, potentially skewing the interpretation of the relationship between variables.]{style="background-color: lightblue;"}

[3. **Contextual Understanding**: Viewing the actual data points on a scatter plot provides a richer understanding of the relationship between variables within the specific context of the dataset. This understanding can be crucial for making informed decisions or drawing meaningful conclusions from the data.]{style="background-color: lightblue;"}

[4. **Communication**: Scatter plots are often used for communication purposes, as they provide a clear and intuitive way to visualize the relationship between variables. They can be included in reports, presentations, or academic papers to effectively convey insights from the data to a broader audience.]{style="background-color: lightblue;"}
