---
title: "DS152 Tutorial Sheet 3"
format: 
  html:
    embed-resources: true
    self-contained-math: true
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
biopsy <- MASS::biopsy

```

## Exercise

A breast cancer study was conducted by Dr William Wolberg from the University of Wisconsin-Madison. He assessed biopsies of breast tumours for 699 patients up to 15 July 1992, and recorded nine different attributes, which are scores based on a scale from 1 to 10 on different features of the tumours. The response variable is whether the tumour was "benign" or "malignant". A `glimpse` of the data is shown below.

```{r, eval = TRUE}
biopsy <- MASS::biopsy
glimpse(biopsy)
```

If you would like to see more information, type this in R:

```{r}
?biopsy
```

Answer the following questions.

**(a)** Is this an *observational* study? Why or why not?

**(b)** Is this a *retrospective* or *prospective* study? Or isn't there enough information to tell? Justify your answer.

**(c)** Below is a function you have not met before called `summarise_if`. What do you think this function is doing when being used with `group_by` as shown below? Run this code. what do you learn from the output?

```{r, message = FALSE}
library(tidyverse)
biopsy %>% group_by(class) %>% summarise_if(is.numeric, mean)
```

**(d)** Note, when you run the code above, the summary for V6 has NAs. Try adding na.rm = TRUE to the `summarise_if` function and see what happens. What do you think this achieves?

**(e)** Suppose you would like to predict, based on the attributes, whether the tumour was benign or malignant. Why would *linear* regression not be suitable in this case?

**(f)** What is the code below doing and why?

```{r, message = FALSE}
biopsy <- biopsy %>% mutate(malignant = as.numeric(class == "malignant"))
```

**(g)** Carry out a multiple logistic regression with this data using the binary response variable called malignant and predictors V1 - V9. What are the values of the coefficients?

**(h)** The predictor `V3` is a scale referring to the uniformity of cell shape. How would an increase of 1 unit in this scale affect the prediction of this model as to whether the tumour is malignant or benign?
