---
title: "DS152 Tutorial Sheet 2"
format: html
editor: visual
---

## Instructions

-   Attempt the questions below before your tutorial in the week beginning 2nd March 2026.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Exercise**

Run the code below to read a dataset called `car_evaluation` into R. Look at the data and the variables that you have to work with. More details about the data are given below.

```{r, message = FALSE}
library(tidyverse)
car_evaluation <- read_csv("https://www.dropbox.com/s/s54upw5jhoe6r4q/car_evaluation.csv?raw=1")
```

There are 1594 cars in this database that are classified as acceptable (acc) or unacceptable (unacc) according to the following attributes:

-   buying_price: buying price (low, medium, high, vhigh)

-   maintenance_cost: ("vhigh", "high", "med", "low")

-   doors: ("2","3","4","5more")

-   persons: capacity in terms of persons to carry (2,4,more)

-   lugboot: the size of luggage boot (small, med, big)

-   safety: estimated safety of the car (low, med, high)

**(a)** Complete the code below to create a bar plot that shows the number of observations within each `buying_price` category. What do you learn from this visualisation?

```{r, eval = FALSE}
ggplot(car_evaluation, aes(??)) +
  geom_??
```

(i) Now add `fill = decision` to the `aes()` of the plot. What do you learn from this visualisation?

(ii) Next facet using the `safety` variable. What do you learn from this visualisation?

**(b)** Run the following code which aims to fit a logistic regression model using `buying_price` and `safety` as predictor variables for the decision. This code will give you an error. What is causing the error?

```{r, eval = FALSE}
fit_logistic <- glm(decision ~ buying_price + safety, 
                    family = "binomial", 
                    data = car_evaluation)
fit_logistic
```

**(c)** Complete the code below to create a variable in the dataset called `decision_binary` with "acc" as the reference group (there's an example of this in your lecture notes).

```{r, eval = FALSE}
car_evaluation <- car_evaluation %>%
  mutate(decision_binary = ??)
```

(i) Now complete the code below to run the logistic regression with `decision_binary` as the outcome. Does this give you an error? What do you note about the coefficients?

```{r, eval = FALSE}
fit_logistic <- glm(??)
fit_logistic
```

(ii) Once you have fit the model, run the following code to obtain the accuracy of the decision classification. (Note: The code is similar to the code you used in Lab 3 but here we use summarise when doing the calculation for `percentage_correct`). Change to `eval = TRUE` when you are ready to run this code. Did your model do a good job?

```{r, eval = FALSE}
threshold <- 0.5

car_predict <- car_evaluation %>%
                   mutate(p_hat = predict(fit_logistic, type = "response") %>% round(2),
                          type_pred = ifelse(p_hat >= threshold, 1,0))
 
accuracy_results <- car_predict %>% summarise(n_correct = sum(type_pred == decision_binary),
                                              n_total = n(),
                                              percentage_correct = n_correct*100/n_total)
accuracy_results
```

(iii) Now modify the summary code above to add a `group_by` function that will breakdown the accuracy results by decision category. Change to `eval = TRUE` when you are ready to run this code. What do you learn from these results?

```{r, eval = FALSE}
accuracy_results <- car_predict %>% 
```
