---
title: "DS152 Tutorial Sheet 4"
format: 
  html:
    embed-resources: true
    self-contained-math: true
editor: visual
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Exercise 

```{r, echo = FALSE, eval = FALSE}
set.seed(24)
barley <- data.frame(dose = gl(3, 9, labels = c("10ppm","100ppm","1000ppm")),
                     nutrient = gl(3, 3, labels = c("nitrogen","phosphate","potassium")),
                     yield = rnorm(27, rep(c(4,5,6,7,8,9,10,8,7), each = 3), .5))
barley <- as_tibble(barley)
```

The `barley` dataset contains data on an experiment set up in a completely randomised design with three replicates and two factors: `nutrient` (with three levels: nitrogen, phosphate and potassium) and `dose` (with three levels: 10ppm, 100ppm and 1000ppm), totaling nine treatments. The `yield` was measured in tonnes per hectare.

You can run the code below to glimpse the dataset:

```{r, include=TRUE, echo = TRUE, message=FALSE}
barley <- read_csv("https://www.dropbox.com/s/4p6ziofmxe2bqcf/barley.csv?raw=1")
glimpse(barley)
```

The table below displays means per each dose:nutrient combination:

```{r, message = FALSE}
barley_summarise <- barley %>%
                      group_by(dose, nutrient) %>%
                      summarise(mean_yield = mean(yield) %>% round(2),
                                n = n()) 
barley_summarise
```


**(a)** Use `pivot_wider` on this table to have separate columns for the nitrogen, phosphate and potassium means per dose group as shown below:

```{r}
barley_summ_wide <- barley_summarise %>%
                      pivot_wider(names_from = "nutrient",
                                  values_from = "mean_yield")

barley_summ_wide
```


**(b)** Which treatment (dose:nutrient combination) seems to be associated with the largest yield?

```{r, fig.height = .5}
barley_summarise %>% 
  arrange(desc(mean_yield))
```



**(c)** Which dose seems to be associated with the largest yield?

```{r, fig.height = .5}
barley_summarise %>% group_by(dose) %>% summarise(mean(mean_yield))
## which is the same as...
barley %>% group_by(dose) %>% summarise(mean(yield))
```

 
**(d)** Which nutrient seems to be associated with the largest yield?

```{r, fig.height = .5}
barley_summarise %>% group_by(nutrient) %>% summarise(mean(mean_yield))
## which is the same as...
barley %>% group_by(nutrient) %>% summarise(mean(yield))
```


**(e)** The figure below displays an 'interaction plot', as seen in class. Does there appear to be an interaction between the factors `dose` and `nutrient`? Why?

```{r, fig.width = 5, fig.height = 3.5, fig.align = "center", message = FALSE}
barley %>%
  group_by(dose, nutrient) %>%
  summarise(mean = mean(yield)) %>%
  ggplot(aes(x = nutrient, y = mean, colour = dose, group = dose)) +
  theme_bw() +
  geom_point() +
  geom_line()
```

